---
title: "Vignette: Penalized Regression Techniques"
author: Aleksander Cichosz, Anni Li, Brian Che, Justin Vo, Noa Rapoport
format: html
editor: visual
execute:
  message: false
  warning: false
  echo: true
  cache: true
---

## Introduction

A vignette on implementing penalized regression models using life expectancy data as a class project for PSTAT197A in Fall 2022.

**Objectives**

In this vignette you'll learn how to implement:

- Ridge Regression
- LASSO Regression
- Elastic Net Regression


**Prerequisites**

Follow the action item below to get set up for the lab. You may need to install one or more packages if the `library()` calls return an error.

::: callout-important
## Action

**Setup**

1.  Create a new script to follow along for the vignette.
2.  Copy-paste the code chunk below at the top of your script to load required packages and data.
:::

```{r, include = FALSE}
# load packages
library(tidyverse)
library(tidymodels)
library(modelr)
library(rsample)
library(yardstick)
```

### About Our Dataset: Life Expectancy (WHO)

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```
You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

## Exploratory Data Analysis

```{r}

```

## Simple Regression

```{r}

```

## Ridge Regression

```{r}

```

## Lasso Regression

```{r}

```

## Elastic Net Regression

```{r}

```

## Conclusion
