---
title: "Vignette: Penalized Regression Techniques"
author: Aleksander Cichosz, Anni Li, Brian Che, Justin Vo, Noa Rapoport
format: html
editor: visual
execute:
  message: false
  warning: false
  echo: true
  cache: true
---

## Introduction

A vignette on implementing penalized regression models using life expectancy data as a class project for PSTAT197A in Fall 2022.

**Objectives**

In this vignette you'll learn how to implement:

-   Ridge Regression
-   LASSO Regression
-   Elastic Net Regression

**Prerequisites**

Follow the action item below to get set up for the lab. You may need to install one or more packages if the `library()` calls return an error.

::: callout-important
## Action

**Setup**

1.  Create a new script to follow along for the vignette.
2.  Copy-paste the code chunk below at the top of your script to load required packages and data.
:::

```{r, include = FALSE}
# Set Up
# load packages
library(tidyverse)
library(tidymodels)
library(modelr)
library(rsample)
library(yardstick)
library(DataExplorer)

# read in csv file
life_clean <- read.csv("data/life_clean.csv")
```

**About Our Dataset: Life Expectancy (WHO)**

## Exploratory Data Analysis

```{r}
# view first few rows of dataset
head(life_clean)

# dimensions of dataset: 2938 rows x 20 columns
dim(life_clean)

```

```{r}
# plots frequency of missing values for each feature
plot_missing(life_clean)
```

## Simple Regression

```{r}

```

## Ridge Regression

```{r}

```

## Lasso Regression

```{r}

```

## Elastic Net Regression

```{r}

```

## Conclusion
